Supplemental Online Materials:\\\textit{Meta-Analysis of the `Ironic' Effects of Intergroup Contact}

### 2021-06-12 manuscript/manuscript.Rmd
revise abstract

---
title: "Meta-Analysis of the 'Ironic' Effects of Intergroup Contact"
header: "'IRONIC' EFFECTS META-ANALYSIS"
keywords: intergroup contact, meta-analysis, perceived discrimination, collective action, policy support
bibliography: references.bib
csl: apa.csl
nocite: | 
  @19, @45, @46, @91, @93, @244, @284, @286, @303, @322, @324, @325, @336, @401, @414, @423, @480, @609, @703, @733, @783, @789, @803, @813, @823, @829, @830, @856, @893, @894, @933, @956, @1020, @1042, @1045, @1114, @1163, @1221, @1248, @1376, @1386, @1395, @1407, @1411, @1412, @1494, @1525, @1549, @1576, @1614, @1658, @1695, @1743, @1762, @1807, @1850, @1891, @1914, @1949, @1966, @1993, @2001, @2023, @2033, @2075, @2257, @2309, @2333, @2341, @2375, @2380, @2381, @2382, @2383, @2385, @2391, @2392, @2394, @2395, @2396, @2397, @2398, @2399, @3000, @3005, @3051, @3054, @3060, @3134, @3196, @3201, @3205, @4000, @4001, @4002, @4003, @4004, @4005
abstract: 
  Growing evidence suggests that intergroup contact, psychology’s most-researched paradigm for reducing prejudice, has the ‘ironic’ effect of reducing support for social change in disadvantaged groups. We conducted a preregistered meta-analytic test of this effect across `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(dplyr::n_distinct(id)))` studies with `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(length(sample)))` samples of `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(sum(n)))` disadvantaged-group members. As predicted, intergroup contact was, on average, associated with less perceived injustice (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "pi"]), digits = 2), "$"))`), collective action (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "ca"]), digits = 2), "$"))`), and support for reparative policies (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "ps"]), digits = 2), "$"))`). However, these associations were small, variable, and consistent with alternative explanations. Moderator analyses found that associations were strongest in studies of adults that measured intergroup contact directly and examined short-term migration or post-colonial intergroup relations. After controlling for the positive association of negative contact with support for social change, positive contact was not associated with any outcome in studies that measured both predictors. We close by highlighting open questions about the relationship between intergroup contact and social change.
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: apa-template.tex
  word_document:
    reference_docx: word-template.docx
---

@@

Another alternative explanation for the apparent 'ironic' effects of intergroup contact centers on intragroup contact. Whereas most studies focused on contact with the advantaged outgroup, some studies examined contact with the disadvantaged ingroup as a predictor of support for social change. @sengupta_ingroup_2015 found ingroup contact to be associated with decreased endorsement of system-justifying beliefs and increased support for reparative policies benefiting the disadvantaged ingroup, suggesting that contact with other ingroup members can motivate disadvantaged-group members to support social change [see also @1695]. As outgroup contact places limits on how much ingroup contact one can have [@pfister_contact_2020], disadvantaged-group members who have more outgroup contact can be expected to have less ingroup contact. Based on these observations, one might argue that the mobilizing effect of ingroup contact could be an alternative explanation for the negative association between outgroup contact and support for social change observed in prior studies.

>>>>

---
title: "Meta-Analysis of the 'Ironic' Effects of Intergroup Contact"
header: "'IRONIC' EFFECTS META-ANALYSIS"
keywords: intergroup contact, meta-analysis, perceived discrimination, collective action, policy support
bibliography: references.bib
csl: apa.csl
nocite: | 
  @19, @45, @46, @91, @93, @244, @284, @286, @303, @322, @324, @325, @336, @401, @414, @423, @480, @609, @703, @733, @783, @789, @803, @813, @823, @829, @830, @856, @893, @894, @933, @956, @1020, @1042, @1045, @1114, @1163, @1221, @1248, @1376, @1386, @1395, @1407, @1411, @1412, @1494, @1525, @1549, @1576, @1614, @1658, @1695, @1743, @1762, @1807, @1850, @1891, @1914, @1949, @1966, @1993, @2001, @2023, @2033, @2075, @2257, @2309, @2333, @2341, @2375, @2380, @2381, @2382, @2383, @2385, @2391, @2392, @2394, @2395, @2396, @2397, @2398, @2399, @3000, @3005, @3051, @3054, @3060, @3134, @3196, @3201, @3205, @4000, @4001, @4002, @4003, @4004, @4005
abstract: 
  Growing evidence suggests that intergroup contact, psychology’s most-researched paradigm for reducing prejudice, has the ‘ironic’ effect of reducing support for social change in disadvantaged groups. We conducted a preregistered meta-analytic test of this effect across `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(dplyr::n_distinct(id)))` studies with `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(length(sample)))` samples of `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(sum(n)))` disadvantaged-group members. As predicted, intergroup contact was, on average, associated with less perceived injustice (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "pi"]), digits = 2), "$"))`), collective action (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "ca"]), digits = 2), "$"))`), and support for reparative policies (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "ps"]), digits = 2), "$"))`). However, these associations were small, variable, and consistent with alternative explanations. Across outcomes, `r with(readRDS("../results/results_for_abstract.rds"), paste0(round(min(p_pred)*100), "--", round(max(p_pred)*100), "\\%"))` of studies found *positive* associations with intergroup contact. Moderator analyses explained about a third of the between-sample variance, showing that, on average, associations were negative only in studies of adults that measured intergroup contact directly and were strongest in studies that examined short-term migration or (post-)colonial intergroup relations. We found evidence for an alternative explanation for the apparent ‘ironic’ effects of intergroup contact as, after controlling for the positive association of negative contact with support for social change, positive contact was not associated with any of the outcomes. We close by discussing strengths and limitations of the available evidence and by highlighting open questions about the relationship between intergroup contact and social change in disadvantaged groups.
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: apa-template.tex
  word_document:
    reference_docx: word-template.docx
---

@@

Another alternative explanation for the apparent 'ironic' effects of intergroup contact centers on intragroup contact. Whereas most studies focused on contact with the advantaged outgroup, some studies examined contact with the disadvantaged ingroup as a predictor of support for social change. @sengupta_ingroup_2015 found ingroup contact to be associated with decreased endorsement of system-justifying beliefs and increased support for reparative policies benefiting the disadvantaged ingroup, suggesting that contact with other ingroup members can motivate disadvantaged-group members to support social change [see also @1695]. As outgroup contact places limits on how much ingroup contact a person can have [@pfister_contact_2020], disadvantaged-group members who have more outgroup contact can be expected to have less ingroup contact. Based on these observations, one might argue that the mobilizing effect of ingroup contact could be an alternative explanation for the negative association between outgroup contact and support for social change observed in prior studies.

### 2021-06-12 manuscript/manuscript.tex
revise abstract

\noindent Growing evidence suggests that intergroup contact, psychology's
most-researched paradigm for reducing prejudice, has the `ironic' effect
of reducing support for social change in disadvantaged groups. We
conducted a preregistered meta-analytic test of this effect across 98
studies with 140 samples of 213,085 disadvantaged-group members. As
predicted, intergroup contact was, on average, associated with less
perceived injustice (\(r = -.07\)), collective action (\(r = -.06\)),
and support for reparative policies (\(r = -.07\)). However, these
associations were small, variable, and consistent with alternative
explanations. Moderator analyses found that associations were strongest
in studies of adults that measured intergroup contact directly and
examined short-term migration or post-colonial intergroup relations.
After controlling for the positive association of negative contact with
support for social change, positive contact was not associated with any
outcome in studies that measured both predictors. We close by
highlighting open questions about the relationship between intergroup
contact and social change.

@@

Another alternative explanation for the apparent `ironic' effects of
intergroup contact centers on intragroup contact. Whereas most studies
focused on contact with the advantaged outgroup, some studies examined
contact with the disadvantaged ingroup as a predictor of support for
social change. Sengupta et al. (2015) found ingroup contact to be
associated with decreased endorsement of system-justifying beliefs and
increased support for reparative policies benefiting the disadvantaged
ingroup, suggesting that contact with other ingroup members can motivate
disadvantaged-group members to support social change (see also Sengupta
\& Sibley, 2013). As outgroup contact places limits on how much ingroup
contact one can have (Pfister et al., 2020), disadvantaged-group members
who have more outgroup contact can be expected to have less ingroup
contact. Based on these observations, one might argue that the
mobilizing effect of ingroup contact could be an alternative explanation
for the negative association between outgroup contact and support for
social change observed in prior studies.

>>>>

\noindent Growing evidence suggests that intergroup contact, psychology's
most-researched paradigm for reducing prejudice, has the `ironic' effect
of reducing support for social change in disadvantaged groups. We
conducted a preregistered meta-analytic test of this effect across 98
studies with 140 samples of 213,085 disadvantaged-group members. As
predicted, intergroup contact was, on average, associated with less
perceived injustice (\(r = -.07\)), collective action (\(r = -.06\)),
and support for reparative policies (\(r = -.07\)). However, these
associations were small, variable, and consistent with alternative
explanations. Across outcomes, 25--36\% of studies found \emph{positive}
associations with intergroup contact. Moderator analyses explained about
a third of the between-sample variance, showing that, on average,
associations were negative only in studies of adults that measured
intergroup contact directly and were strongest in studies that examined
short-term migration or (post-)colonial intergroup relations. We found
evidence for an alternative explanation for the apparent `ironic'
effects of intergroup contact as, after controlling for the positive
association of negative contact with support for social change, positive
contact was not associated with any of the outcomes. We close by
discussing strengths and limitations of the available evidence and by
highlighting open questions about the relationship between intergroup
contact and social change in disadvantaged groups.

@@

Another alternative explanation for the apparent `ironic' effects of
intergroup contact centers on intragroup contact. Whereas most studies
focused on contact with the advantaged outgroup, some studies examined
contact with the disadvantaged ingroup as a predictor of support for
social change. Sengupta et al. (2015) found ingroup contact to be
associated with decreased endorsement of system-justifying beliefs and
increased support for reparative policies benefiting the disadvantaged
ingroup, suggesting that contact with other ingroup members can motivate
disadvantaged-group members to support social change (see also Sengupta
\& Sibley, 2013). As outgroup contact places limits on how much ingroup
contact a person can have (Pfister et al., 2020), disadvantaged-group
members who have more outgroup contact can be expected to have less
ingroup contact. Based on these observations, one might argue that the
mobilizing effect of ingroup contact could be an alternative explanation
for the negative association between outgroup contact and support for
social change observed in prior studies.

### 2021-06-12 manuscript/supplemental-materials.Rmd
revise abstract

First, we ran a series of Bayesian random-effects meta-regression models to estimate how much of the between-samples heterogeneity was explained by specific categorical moderator variables (`run_moderator_analyses.R`). Figures S1 and S2 show the results for collective action and policy support. Comparisons were inconclusive because we had an insufficient number of effect sizes per category. Second, we ran another three random-effects meta-regression models with Muthukrishna et al.'s [-@muthukrishna_beyond_2020] measure of cultural distance, where available, as a continuous moderator variable. Because cultural distance was a country-level moderator variable, we used two-level random-effects models in which we estimated country-specific deviations from the mean effect size as well as sample-specific deviations from the country-specific effect size. Figure 5 (in the main text) show the inconclusive results for all three outcomes.

Second, we used meta-regression trees to discover interactions between moderator variables that best explained heterogeneity in effect sizes [@li_meta-cart_2017]. As recommended, we ran random-effects meta-regression tree analyses using the look-ahead strategy [@li_multiple_2020]. We set the pruning parameter to $c = 0$ because, for this analysis, we prioritized exploration over error control (`run_exploratory_moderator_analyses.R`). For both collective action and policy support, the algorithm found that no moderator or interaction of moderators explained between-samples heterogeneity. This result is not surprising, however, as both analyses used fewer than 40 effect sizes, the minimum number required for the algorithm to perform well in detecting even simple interaction effects [@li_meta-cart_2017]. As all moderator analyses were inconclusive for collective action and policy support, we only report results for perceived injustice in the main text.

>>>>

First, we ran a series of Bayesian random-effects meta-regression models to estimate how much of the between-samples heterogeneity was explained by specific categorical moderator variables (`run_moderator_analyses.R`). Figures S1 and S2 show the results for collective action and policy support. Comparisons were inconclusive because we had an insufficient number of effect sizes per category.

Second, we ran another three random-effects meta-regression models with Muthukrishna et al.'s [-@muthukrishna_beyond_2020] measure of cultural distance, where available, as a continuous moderator variable. Because cultural distance was a country-level moderator variable, we used two-level random-effects models in which we estimated country-specific deviations from the mean effect size as well as sample-specific deviations from the country-specific effect size. Figure 5 (in the main text) show the inconclusive results for all three outcomes.

Third, we used meta-regression trees to discover interactions between moderator variables that best explained heterogeneity in effect sizes [@li_meta-cart_2017]. As recommended, we ran random-effects meta-regression tree analyses using the look-ahead strategy [@li_multiple_2020]. We set the pruning parameter to $c = 0$ because, for this analysis, we prioritized exploration over error control (`run_exploratory_moderator_analyses.R`). For both collective action and policy support, the algorithm found that no moderator or interaction of moderators explained between-samples heterogeneity. This result is not surprising, however, as both analyses used fewer than 40 effect sizes, the minimum number required for the algorithm to perform well in detecting even simple interaction effects [@li_meta-cart_2017]. As all moderator analyses were inconclusive for collective action and policy support, we only report results for perceived injustice in the main text.

### 2021-06-12 manuscript/supplemental-materials.tex
revise abstract

\hypertarget{robustness-checks}{%
\subsection{Robustness Checks}\label{robustness-checks}}

As preregistered, we conducted two kinds of robustness checks. First, we
assessed to what extent our findings were sensitive to choosing
narrower, \(\mu \sim \text{Normal}(0, 0.1)\), or wider,
\(\mu \sim \text{Normal}(0, 1)\), prior distributions. Choosing narrower
or wider prior distribution did not affect mean effect size estimates
for perceived injustice (\(\Delta r = -.00, [-.05, .04]\) and
\(\Delta r = .00, [-.05, .05]\)), collective action
(\(\Delta r = -.01, [-.11, .09]\) and \(\Delta r = .00, [-.11, .10]\)),
and policy support (\(\Delta r = -.01, [-.10, .08]\) and
\(\Delta r = .00, [-.09, .10]\)). Second, we assessed to what extent our
findings were sensitive to including or excluding influential studies by
repeating the preregistered analyses \(J\) times while leaving out one
of \(J\) studies each time and by calculating the mean absolute
difference (\emph{MAD}) for the estimated mean effect size across
left-out studies. For perceived injustice
(\(\textit{MAD} = .02, [.01, .04]\)), collective action
(\(\textit{MAD} = .04, [.02, .09]\)), and policy support
(\(\textit{MAD} = .03, [.02, .08]\)), the \emph{MAD} was small. Leaving
out the most influential study, for example, did not change estimates of
the mean effect size for the three outcomes
(\(\Delta r = .01, [-.04, .05]\); \(\Delta r = .02, [-.08, .12]\);
\(\Delta r = -.02, [-.11, .07]\)). Together, these analyses showed that
our findings were robust to choosing different prior distributions and
to excluding influential studies.

@@

First, we ran a series of Bayesian random-effects meta-regression models
to estimate how much of the between-samples heterogeneity was explained
by specific categorical moderator variables
(\texttt{run\_moderator\_analyses.R}). Figures S1 and S2 show the
results for collective action and policy support. Comparisons were
inconclusive because we had an insufficient number of effect sizes per
category. We ran another three random-effects meta-regression models
with Muthukrishna et al.'s (2020) measure of cultural distance, where
available, as a continuous moderator variable. Because cultural distance
was a country-level moderator variable, we used two-level random-effects
models in which we estimated country-specific deviations from the mean
effect size as well as sample-specific deviations from the
country-specific effect size. Figure 5 (in the main text) show the
inconclusive results for all three outcomes.

Second, we used meta-regression trees to discover interactions between
moderator variables that best explained heterogeneity in effect sizes
(Li et al., 2017, 2020). As recommended, we ran random-effects
meta-regression tree analyses using the look-ahead strategy (Li et al.,
2020). We set the pruning parameter to \(c = 0\) because, for this
analysis, we prioritized exploration over error control
(\texttt{run\_exploratory\_moderator\_analyses.R}). For both collective
action and policy support, the algorithm found that no moderator or
interaction of moderators explained between-samples heterogeneity. This
result is not surprising, however, as both analyses used fewer than 40
effect sizes, the minimum number required for the algorithm to perform
well in detecting even simple interaction effects (Li et al., 2017). As
all moderator analyses were inconclusive for collective action and
policy support, we only report results for perceived injustice in the
main text.

>>>>

@@

First, we ran a series of Bayesian random-effects meta-regression models
to estimate how much of the between-samples heterogeneity was explained
by specific categorical moderator variables
(\texttt{run\_moderator\_analyses.R}). Figures S1 and S2 show the
results for collective action and policy support. Comparisons were
inconclusive because we had an insufficient number of effect sizes per
category.

Second, we ran another three random-effects meta-regression models with
Muthukrishna et al.'s (2020) measure of cultural distance, where
available, as a continuous moderator variable. Because cultural distance
was a country-level moderator variable, we used two-level random-effects
models in which we estimated country-specific deviations from the mean
effect size as well as sample-specific deviations from the
country-specific effect size. Figure 5 (in the main text) show the
inconclusive results for all three outcomes.

Third, we used meta-regression trees to discover interactions between
moderator variables that best explained heterogeneity in effect sizes
(Li et al., 2017). As recommended, we ran random-effects meta-regression
tree analyses using the look-ahead strategy (Li et al., 2020). We set
the pruning parameter to \(c = 0\) because, for this analysis, we
prioritized exploration over error control
(\texttt{run\_exploratory\_moderator\_analyses.R}). For both collective
action and policy support, the algorithm found that no moderator or
interaction of moderators explained between-samples heterogeneity. This
result is not surprising, however, as both analyses used fewer than 40
effect sizes, the minimum number required for the algorithm to perform
well in detecting even simple interaction effects (Li et al., 2017). As
all moderator analyses were inconclusive for collective action and
policy support, we only report results for perceived injustice in the
main text.
